### 2017-02-01 local-capability-machine.tex
Latex fix, add note to RW.
Fix another merge artifact.
Fix the last fix.
Some more work on the FTLR proof.

@@

\begin{lemma}[Conditions for load instruction are sufficient]
  \label{lem:conds-load-suffice}
  \item $\heapSat[\memheap]{n}{W}$
  \item $\var{c} = ((\perm,\gl),\start,\addrend,\addr)$
  \item $\npair{c}\in\stdvr(W)$
  \item $\readAllowed{\perm}$ 
  \item $\withinBounds{\var{c}}$
  then $\npair[n-1]{\memheap(\addr)} \in \stdvr(W)$
\end{lemma}
\begin{proof}
  \begin{enumproof}
  \item $\npair{(\start,\addrend)} \in \readCond{\gl}(W)$: follows by
    definition of $\stdvr{\cdot}$ from $\npair{c}\in\stdvr(W)$.
  \item $\exists r \in \var{localityReg}(g,W)$, $[\start',\addrend'] \supseteq
    [\start,\addrend] \ldotp$ $W(r)\nsubsim[n-1]
    \iota_{\start',\addrend'}^\pwl$. By definition of $\readCond{\gl}(W)$. \label{step:unfold-readCond}
  \item $\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp
    \memSatPar{\memheap}{W}{P}$. By definition of $\heapSat[\memheap]{n}{W}$.
  \item $\memheap = \biguplus_{r\in\activeReg{W}}P(r)$ and $\forall r \in
    \activeReg{W}$, $n'<n \ldotp$ $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
    and $\npair[n']{P(r)} \in H(s)(\xi^{-1}(W))$. By definition of
    $\memSatPar{\memheap}{W}{P}$. \label{step:unfold-phi-sat-part}
  \item $r \in \var{localityReg}(g,W)\subseteq \activeReg{W}$. By definition of
    $\var{localityReg}(\cdot)$ and $\activeReg{\cdot}$.
  \item $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
    and $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$.
    By specializing the result from
    Step~\ref{step:unfold-phi-sat-part} to the $r$ from
    Step~\ref{step:unfold-readCond} and $n' = n-1$.
  \item $\npair[n-1]{P(r)} \in H^\pwl_{\start',\addrend'}(s)(\xi^{-1}(W))$.
    Follows by combining $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$ with
    $W(r)\nsubsim[n-1] \iota_{\start',\addrend'}^\pwl$ from Step~\ref{step:unfold-readCond}.
  \item $\dom(P(r)) = [\start',\addrend']$ and for all $\addr' \in
    [\start',\addrend']\ldotp$
    $\npair[n-2]{P(r)(\addr')}\in\stdvr(\xi(\xi^{-1}(W)))$.
  \item $\addr \in [\start,\addrend]\subseteq[\start',\addrend']$. By combining
    $\withinBounds{c}$ with the fact that $[\start',\addrend'] \supseteq
    [\start,\addrend] \ldotp$ from Step~\ref{step:unfold-readCond}.
  \item In particular, we get: $\memheap(\addr) = P(r)(\addr)$ and
    $\npair[n-2]{P(r)(\addr)}\in\stdvr(W)$.
    \dominique{TODO: Make sure we lose only one step in the definitions!}
  \end{enumproof}
\end{proof}

@@

  \item $n > 0$, since otherwise $i = 0$ (because $i \leq n$) and this is

@@

  \item $\npair{\memreg(\pcreg)}\in\stdvr(W)$. \label{step:ftlr-pc-vr}

@@

      Suffices: $\npair{(\perm',\start,\addrend)}\in\execCond{g}(W)$\\

@@

    \item Assume: $n'<n$, $W'\future W$, $\addr' \in [\start,\addrend]$, $\gl

@@

      Suffices: $\npair[n']{((\perm,\gl),\start,\addrend,\addr')} \in

@@

  \item For all $r \in \RegName$, $\npair[n]{\memreg(r)}\in\stdvr(W)$. \label{step:ftlr-reg-rr}
    \item Case $r \neq \pcreg$: follows from $\npair[n]{\reg}\in \stdrr(W)$ by

@@

      $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{lem:heap-sat-dc}).
  \item Case $\var{conf'} = \updateReg[\Phi']{\pcreg}{\var{newPc}}$, and additionally, one of the following

@@

    \item $\memheap[\Phi'] = \memheap[\Phi]$
    \item $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$, with

@@

      $\memreg[\Phi'](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$

@@

    \item $\memreg[\Phi'](r) = \memreg[\Phi](r)$
    \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$
    \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =

@@

    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =

@@

      $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + n)$ for some $n \in

@@

    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =

@@

    \item $\heapSat[{\memheap[\Phi']}]{n-1}{W}$.
      \item Case $\memheap[\Phi'] = \memheap[\Phi]$:
        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
      \item Case $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$,

@@

        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by

@@

        fact that $\heapSat[\memheap]{n}{W}$ and
        $\npair{\memreg(r_1)}\in\stdvr(W)$ and
        $\npair{\memreg(r_2)}\in\stdvr(W)$ which follows from
        $\npair{\memreg}\in\stdrr(W)$ by definition.
    \item For all $r \in \RegName$, $\npair[n-1]{\memreg[\Phi'](r)}\in\stdvr(W)$.
      \item Case $\memreg[\Phi'](r) = \memreg[\Phi](r)$:
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows from

@@

      \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$.
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by definition of

@@

      \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =

@@

        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by

@@

        $\heapSat[\memheap]{n}{W}$ and $\npair{\memreg(r_2)}\in\stdvr(W)$ which

@@

        fact that $\npair{\memreg(r_1)}\in\stdvr(W)$

@@

        fact that $\npair{\memreg(r)}\in\stdvr(W)$
        which follows from $\npair{\memreg}\in\stdrr(W)$ by

@@

        fact that $\npair{\memreg(r)}\in\stdvr(W)$
        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
    \item $\npair[n-1]{\memreg[\Phi']}\in\stdrr(W)$: Follows from the previous

@@

    \item $\npair[n-1]{\var{newPc}} \in \stder(W)$:

@@

          \item $\npair[n-1]{\memreg(\lv)} \in \stdvr(W)$.  Follows from
          \item $\npair[n-1]{(\start',\addrend',addr') \in
              \entryCond{\gl'}(W)}$. By definition of $\stdvr(W)$ from the

@@

@@

        $\memreg[\Phi'](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$: TODO

@@

\cite{Devriese:2016ObjCap} describe a logical relation for a JavaScript-like
language with object capabilities.  Relevant because it treats object
capabilities, albeit in a JavaScript-like lambda calculus.

>>>>

\begin{lemma}[Failing terms are in $\observations$ and $\stder$]
  \label{lem:failed-obs-stder}
  If $(\reg,\ms\uplus\ms_f) \step[*] \failed$ for all $\ms_f$, then
  $\npair{(\reg,\ms)} \in \observations(W)$ for any $W$.

  If $(\reg\update{\pcreg}{\word},\ms) \step[*] \failed$ for all $\reg, \ms$,
  then $\npair{\word} \in \stder(W)$ for any $W$.
\end{lemma}
\begin{proof}
  Follows from the definitions of $\observations(W)$ and $\stder(W)$ using an
  (omitted) determinacy result. \dominique{at least state the result
    somewhere?}
\end{proof}

@@

\begin{lemma}[Conditions for load instruction are sufficient]
  \label{lem:conds-load-suffice}
  If
  \begin{itemize}
  \item $\heapSat[\memheap]{n}{W}$
  \item $\var{c} = ((\perm,\gl),\start,\addrend,\addr)$
  \item $\npair{c}\in\stdvr(W)$
  \item $\readAllowed{\perm}$
  \item $\withinBounds{\var{c}}$
  \end{itemize}
  then $\npair[n-1]{\memheap(\addr)} \in \stdvr(W)$
\end{lemma}
\begin{proof}
  \begin{enumproof}
  \item $\npair{(\start,\addrend)} \in \readCond{\gl}(W)$: follows by
    definition of $\stdvr{\cdot}$ from $\npair{c}\in\stdvr(W)$.
  \item $\exists r \in \var{localityReg}(g,W)$, $[\start',\addrend'] \supseteq
    [\start,\addrend] \ldotp$ $W(r)\nsubsim[n-1]
    \iota_{\start',\addrend'}^\pwl$. By definition of $\readCond{\gl}(W)$. \label{step:unfold-readCond}
  \item $\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp
    \memSatPar{\memheap}{W}{P}$. By definition of $\heapSat[\memheap]{n}{W}$.
  \item $\memheap = \biguplus_{r\in\activeReg{W}}P(r)$ and $\forall r \in
    \activeReg{W}$, $n'<n \ldotp$ $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
    and $\npair[n']{P(r)} \in H(s)(\xi^{-1}(W))$. By definition of
    $\memSatPar{\memheap}{W}{P}$. \label{step:unfold-phi-sat-part}
  \item $r \in \var{localityReg}(g,W)\subseteq \activeReg{W}$. By definition of
    $\var{localityReg}(\cdot)$ and $\activeReg{\cdot}$.
  \item $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
    and $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$.
    By specializing the result from
    Step~\ref{step:unfold-phi-sat-part} to the $r$ from
    Step~\ref{step:unfold-readCond} and $n' = n-1$.
  \item $\npair[n-1]{P(r)} \in H^\pwl_{\start',\addrend'}(s)(\xi^{-1}(W))$.
    Follows by combining $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$ with
    $W(r)\nsubsim[n-1] \iota_{\start',\addrend'}^\pwl$ from Step~\ref{step:unfold-readCond}.
  \item $\dom(P(r)) = [\start',\addrend']$ and for all $\addr' \in
    [\start',\addrend']\ldotp$
    $\npair[n-2]{P(r)(\addr')}\in\stdvr(\xi(\xi^{-1}(W)))$.
  \item $\addr \in [\start,\addrend]\subseteq[\start',\addrend']$. By combining
    $\withinBounds{c}$ with the fact that $[\start',\addrend'] \supseteq
    [\start,\addrend] \ldotp$ from Step~\ref{step:unfold-readCond}.
  \item In particular, we get: $\memheap(\addr) = P(r)(\addr)$ and
    $\npair[n-2]{P(r)(\addr)}\in\stdvr(W)$.
    \dominique{TODO: Make sure we lose only one step in the definitions!}
  \end{enumproof}
\end{proof}

@@

    \end{equation*} \label{step:ip-non-zero}
  \item $n' > 0$, since otherwise $i = 0$ (because $i \leq n'\leq n$) and this is

@@

  \item $\npair[n']{\memreg(\pcreg)}\in\stdvr(W)$. \label{step:ftlr-pc-vr}

@@

      Suffices: $\npair[n']{(\perm',\start,\addrend)}\in\execCond{g}(W)$\\

@@

    \item Assume: $n''<n'$, $W'\future W$, $\addr' \in [\start,\addrend]$, $\gl

@@

      Suffices: $\npair[n'']{((\perm,\gl),\start,\addrend,\addr')} \in

@@

  \item For all $r \in \RegName$, $\npair[n']{\memreg(r)}\in\stdvr(W)$. \label{step:ftlr-reg-rr}
    \item Case $r \neq \pcreg$: follows from $\npair[n']{\reg}\in \stdrr(W)$ by

@@

      $\heapSat[\hs']{n'-1}{W'}$ (using Lemma~\ref{lem:heap-sat-dc}).
  \item Case $\var{conf'} = \updateReg[\Phi'']{\pcreg}{\var{newPc}}$, and additionally, one of the following

@@

    \item $\memheap[\Phi''] = \memheap[\Phi]$
    \item $\memheap[\Phi''] = \memheap[\Phi]\update{\addr'}{\var{w}}$, with

@@

      $\memreg[\Phi''](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$

@@

    \item $\memreg[\Phi''](r) = \memreg[\Phi](r)$
    \item $\memreg[\Phi''](r) = z$ for some $z \in \ints$
    \item $\memreg[\Phi''](r) = \var{w}$ and $\memreg(r_2) =

@@

    \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r_1) =

@@

      $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + z)$ for some $z \in

@@

    \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r) =

@@

    \item $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$.
      \item Case $\memheap[\Phi''] = \memheap[\Phi]$:
        $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
      \item Case $\memheap[\Phi''] = \memheap[\Phi]\update{\addr'}{\var{w}}$,

@@

        $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$ follows by

@@

        fact that $\heapSat[\memheap]{n'}{W}$ and
        $\npair[n']{\memreg(r_1)}\in\stdvr(W)$ and
        $\npair[n']{\memreg(r_2)}\in\stdvr(W)$ which follows from
        $\npair[n']{\memreg}\in\stdrr(W)$ by definition.
    \item For all $r \in \RegName$, $\npair[n'-1]{\memreg[\Phi''](r)}\in\stdvr(W)$.\label{step:phip-regs-stdvr}
      \item Case $\memreg[\Phi''](r) = \memreg[\Phi](r)$:
        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows from

@@

      \item $\memreg[\Phi''](r) = z$ for some $z \in \ints$.
        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by definition of

@@

      \item $\memreg[\Phi''](r) = \var{w}$ and $\memreg(r_2) =

@@

        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by

@@

        $\heapSat[\memheap]{n'}{W}$ and $\npair[n']{\memreg(r_2)}\in\stdvr(W)$ which

@@

        fact that $\npair[n']{\memreg(r_1)}\in\stdvr(W)$

@@

        fact that $\npair[n']{\memreg(r)}\in\stdvr(W)$
        which follows from $\npair[n']{\memreg}\in\stdrr(W)$ by

@@

        fact that $\npair[n']{\memreg(r)}\in\stdvr(W)$
        which follows from $\npair[n']{\memreg}\in\stdrr(W)$ by definition.
    \item $\npair[n'-1]{\memreg[\Phi'']}\in\stdrr(W)$: Follows from the previous

@@

    \item $\npair[n'-1]{\var{newPc}} \in \stder(W)$:

@@

          \item $\npair[n']{\memreg(\lv)} \in \stdvr(W)$.  Follows from
            Step~\ref{step:ftlr-reg-rr}.
          \item $\npair[n']{(\start',\addrend',addr')} \in \entryCond{\gl'}(W)$.
            By definition of $\stdvr(W)$ from the previous point.
          \item $\npair[n'-1]{((\exec,\gl'),\start',\addrend',\addr')} \in
            \stder(W)$: By definition of $\entryCond{\cdot}$ and taking $n' =
            n'-1$ and $W' = W$
          \item $\updatePcPerm{\memreg(\lv)} =
            ((\exec,\gl'),\start',\addrend',\addr')$: by definition of $\updatePcPerm{\cdot}$.
        \item Case $\memreg(\lv) = ((\perm',\gl'),\start',\addrend',\addr')$
          with $\perm' \in \{\exec,\rwx,\rwlx\}$ and $\withinBounds{\memreg(\lv)}$:
          \item $\npair[n']{\memreg(\lv)} \in \stdvr(W)$.  Follows from
            Step~\ref{step:ftlr-reg-rr}.
          \item $\npair[n']{(\perm',\start',\addrend',\addr')} \in \execCond{\gl'}(W)$.
            By definition of $\stdvr(W)$ from the previous point.
          \item $\npair[n'-1]{((\perm',\gl'),\start',\addrend',\addr')} \in
            \stder(W)$: By definition of $\execCond{\cdot}$, taking $n' = n'-1$,
            $W' = W$ and $\addr = \addr'$. Note that $\addr' \in
            [\start',\addrend']$ because we have $\withinBounds{\memreg(\lv)}$.
          \item $\updatePcPerm{\memreg(\lv)} =
            ((\perm',\gl'),\start',\addrend',\addr')$: by definition of
            $\updatePcPerm{\cdot}$.
        \item Case not ($\memreg(\lv) =
          ((\entry,\gl'),\start',\addrend',\addr')$) and not ($\memreg(\lv) =
          ((\perm',\gl'),\start',\addrend',\addr')$ with $\perm' \in
          \{\exec,\rwx,\rwlx\}$ and $\withinBounds{\memreg(\lv)}$):
          \item $\updatePcPerm{\memreg(\lv)} = \memreg(\lv)$: by definition of 
            $\updatePcPerm{\cdot}$.
          \item $(\reg\update{\pcreg}{\memreg(\lv)},\ms)\step\failed$ for any
            $\reg$,$\ms$: by definition of the evaluation relation.
          \item $\npair[n'-1]{\var{newPc}} \in \stder(W)$: by
            Lemma~\ref{lem:failed-obs-stder} using the previous point.
        $\memreg[\Phi''](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$:
        \item Case $\perm' \in \{\exec,\rwx,\rwlx\}$ and $\start'\leq
          \addr'+1\leq \addrend'$:
          \item $\npair[n'-1]{\memreg[\Phi''](\pcreg)} \in \stdvr(W)$: by
            Step~\ref{step:phip-regs-stdvr}.
          \item $\npair[n'-1]{((\perm',\gl'),\start',\addrend',\addr' + 1)} \in \stdvr(W)$: by
            Lemma~\ref{lem:conds-lea-suffice} from the previous point.
          \item One of the following holds:
                  \perm' = \exec \land\\
                  \npair[n'-1]{(\start',\addrend')} \in \readCond{}(\gl)(W)
                  \perm' = \rwx \land \\
                  \npair[n'-1]{(\start',\addrend')} \in \readCond{}(\gl)(W) \land\\
                  \npair[n'-1]{(\start',\addrend')} \in \writeCond{}(\iota^\nwl,\gl)(W)
                  \perm' = \rwlx \land\\
                  \npair[n'-1]{(\start',\addrend')} \in \readCond{}(\gl)(W) \land\\
                  \npair[n'-1]{(\start',\addrend')} \in \writeCond{}(\iota^\pwl,\gl)(W),
           This follows from the previous point by  definition of $\stdvr(W)$
         \item $\npair[n'-1]{((\perm',\gl'),\start',\addrend',\addr' + 1)} \in
           \stder(W)$: By the induction hypothesis of this lemma using the

@@

        \item Case not ($\perm' \in \{\exec,\rwx,\rwlx\}$ and $\start'\leq
          \addr'+1\leq \addrend'$):
          The result follows by Lemma~\ref{lem:failed-obs-stder}.

@@

      $\npair[n'-1]{(\memreg[\Phi'']\update{\pcreg}{\var{newPc}},\memheap[\Phi''])}\in\observations(W)$,
      i.e. $\npair[n'-1]{\var{conf'}}\in\observations(W)$:
      by definition of $\stder(W)$ using the above three points.
    \item $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
      By: definition of $\observations(W)$ using the previous step and the
      evaluation $\var{conf'}\step[i'] (\halted,\Phi')$ from
      Step~\ref{step:ip-non-zero}.\dominique{TODO: check.  What about ms_f and other stuff.}

@@

\cite{Devriese:2016ObjCap} describe a logical relation for a JavaScript-like
language with object capabilities.  Relevant because it treats object
capabilities, albeit in a JavaScript-like lambda calculus. It also deals with an
untyped language, using a semantic unitype.
